---
title: "U.S. Caribbean \\linebreak Snapshot Ecosystem Status Report"
format:
  pdf:
    margin-left: "0.5in"
    margin-right: "0.5in"
    margin-top: "0.2in"
    margin-bottom: "0.5in"
    fontsize: "10pt"
    linestretch: 1
    indent: true
    include-in-header: tex_template.tex
    keep-tex: true
    papersize: letter
params:
  widths: !expr c(0.9, 0.75, 2, 3) 
  img1: !expr here::here("images/process_flow_chart.png")
  img2: !expr here::here("images/traffic.png")
  img3: !expr here::here("images/pcas.png")
  img_dir: !expr here::here("images")
latex_engine: xelatex
bibliography: bsb_references.bib
nocite: |
  @jeanmichel_copernicus_2021, @fratantoni_description_2015
csl: cambridge-university-press-numeric.csl
---

```{r setup, include=FALSE}
library(here) 
library(flextable) 
library(ggplot2) 
```

\AddToShipoutPictureBG*{%
  \AtPageLowerLeft{%
    \includegraphics[width=\paperwidth,height=\paperheight]{bg_pg1.jpg}%
  }%
}

\vspace{-2.5cm}
\section{2025}

This is a short-form update to the full 2025 U.S. Caribbean Ecosystem Status Report (ESR) [@tabandera_black_2024] highlighting the recent status of environmental, ecological, and socioeconomic factors. Indicators were compiled into two categories: tracking performance toward fishery management objectives and risks to meeting fishery management objectives.

\vspace{0.2cm}

::: {layout="[ [0.4, 0.025, 0.575], [1], [0.3, 0.025, 0.675]] "}

:::: {}
![](`r params$img1`)
\normalsize
::::

:::: {}
\hfill
::::

:::: {}
\section{Overview of recent trends}
\vspace{-0.25cm}

\subsection{Performace indicators}

-   17 indicators were compiled to track perfomance towards management objectives.
-   Indicators were categorized as relating to food production, socioeconomic health, equality, engagement and participation, bycatch reduction, governance, and protection of ecosystems.
-   add some text about indicators

\vspace{-0.25cm}
\subsection{Risk indicators}

-   13 indicators were compiled to track risks to meeting management objectives.
-   Indicators track changes in the physical environment and human activities.
-   Major recent changes in the physical environment include increased sea surface temperature, coral bleaching stress, and ocean acidification. 
-   Other insights?
::::

:::: {}
\vspace{0.5cm}
\section{Integrated Ecosystem Perspectives}
\vspace{-0.25cm}
::::

:::: {}
\raggedright

\subsection{Analysis}
Multivariate methods (principal components analysis and traffic light plot; see details in full report) were used to synthesize the information contained in the full suite of indicators.
\vspace{0.1cm}

\subsection{Interpretation}
The traffic plot conveys that many indicator values underwent rapid change in the period 2017-2021, and the PCA biplots confirm these patterns as there are larger two-dimensional shifts between these years. These shifts are most likely driven by several major stressor events in this time period, including the major hurricanes Maria and Irma (2017) and the COVID pandemic (2020-2021). Together, the multivariate analyses suggest that these events have had some destabilizing impacts on the U.S. Caribbean fishery ecosystem.

::::

:::: {}
\hfill
::::

:::: {}
![](`r params$img2`) 
::::

:::


\newpage

\AddToShipoutPictureBG*{%
  \AtPageLowerLeft{%
    \includegraphics[width=\paperwidth,height=\paperheight]{bg_pg2.jpg}%
  }%
}


\newgeometry{top=0.75in, left=0.75in, right=0.75in, bottom=0.75in}

```{r, sum_table, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(gt)
library(stringr)

sum_table <- readRDS(here::here("docs/indicator_summary_table.rds"))

sum_table <- sum_table %>% 
  filter(`Trend symbol` != "●")

gt_table <- sum_table %>%
  mutate(
    Indicator = str_wrap(Indicator, width = 35)  # Adjust to get desired line length
  ) %>%
  gt() %>%
  tab_row_group(
    label = "Recent trend below average",
    rows = `Trend symbol` == "-"
  ) %>%
  tab_row_group(
    label = "Recent trend above average",
    rows = `Trend symbol` == "+"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  cols_label(
    Indicator = "Indicator"  
  ) %>%
  tab_style(
    style = cell_borders(
      sides = "right",
      color = "light gray",
      weight = px(2)
    ),
    locations = cells_body(
      columns = c(Indicator)
    )
  ) %>%
  cols_width(
  Indicator ~ px(180),         # Column 1 — wide for wrapped text
  Units ~ px(160),     # Column 2
  `Extent/sub-indicator` ~ px(100),           # Column 3
  `Trend symbol` ~ px(40),       # Column 4
  `Slope symbol` ~ px(40),       # Column 5
  Mean ~ px(40),           # Column 6
  SD ~ px(40),             # Column 7
  `Min year` ~ px(40),           # Column 6
  `Max year` ~ px(40)             # Column 7
) %>%
  cols_align(
    align = "left",
    columns = everything()
  ) %>%
  tab_options(
    row_group.background.color = "lightgrey",
    table.font.names = "Times New Roman",
    table.font.size = 10,
    table.width = pct(100)
  )

gt_table
```


\vspace{-0.4cm}
\footnotesize \* The table above shows details for indicators with recent trends (last 5 years of available data) either above or below the average of the time series. For each indicator, the units and extent or sub-indicator name are presented, as well as a symbol for whether the recent trend is above or below average (+ or -) and a symbol for whether the slope of a linear model fit to the last 5 years of data is increasing, decreasing, or stable compared to the long-term slope. The mean of the time series, standard deviation, minimum year and maximum year are also presented. Indicators were excluded if there was no data available after 2020. The code used to create this report can be viewed online: [github.com/NEFSC/READ-EDAB-bsbESP](github.com/NEFSC/READ-EDAB/bsbESP) \newline

\centering\normalsize

We welcome your observations! Please contact [mandy.karnauskas@noaa.gov](mandy.karnauskas@noaa.gov) or [carissa.gervasi@noaa.gov](carissa.gervasi@noaa.gov) with questions or comments on the information presented in this report.


# References
\begingroup
\footnotesize
::: {#refs}
:::
\endgroup

